\LARGE \bf Matrix Multiplication-Driven Repulsive Fields for 3D Voxel-Based Robotic Manipulator Path Planning

### 2024-02-29 Articles/IROS/root.tex
rough draft of convolution and interpolation

% The following packages can be found on http:\\www.ctan.org
%\usepackage{graphics} % for pdf, bitmapped graphics files
%\usepackage{epsfig} % for postscript graphics files
%\usepackage{mathptmx} % assumes new font selection scheme installed
%\usepackage{times} % assumes new font selection scheme installed
%\usepackage{amsmath} % assumes amsmath package installed
%\usepackage{amssymb}  % assumes amsmath package installed
\usepackage{hyperref}

@@

Since the obstacle space is discrete (has finite resolution), while the Cartesian space is continuous, we propose two methods for mapping from Cartesian space to the occupancy grid space. The simpler approach involves mapping the point directly to the center of the nearest occupancy grid voxel, based on Euclidean distance. However, this discretization can sometimes lead to discontinuities. Therefore, we propose a second approach: tri-linear interpolation of the calculated repulsive field to achieve a continuous repulsive field value.

\subsection{AREA SELECTION}

@@

@@

>>>>

% The following packages can be found on http:\\www.ctan.org
%\usepackage{graphics} % for pdf, bitmapped graphics files
%\usepackage{epsfig} % for postscript graphics files
%\usepackage{mathptmx} % assumes new font selection scheme installed
%\usepackage{times} % assumes new font selection scheme installed
\usepackage{amsmath} % assumes amsmath package installed
\usepackage{amssymb}  % assumes amsmath package installed
\usepackage{hyperref}

@@

\subsection{MAPPING}

Since the obstacle space is discrete (has finite resolution), while the Cartesian space is continuous, we propose two methods for mapping from Cartesian space to the occupancy grid space. The simpler approach involves mapping the point directly to the center of the nearest occupancy grid voxel, based on Euclidean distance.\add{is this clear, do i need equation?} However, this discretization can sometimes lead to discontinuities. Therefore, we propose a second approach: tri-linear interpolation of the calculated repulsive field to achieve a continuous repulsive field value.

Once the mapping of these points to the voxel grid is completed, we proceed to employ a specialized kernel convolution method. This method is tailored to calculate the components of the avoidance velocity vector in the Cartesian space, we generate the corresponding kernel and extract a segment of the obstacle grid of matching size, centered at agent or the point of interest (POI), resulting in two same size 3D matrices—one is a "window" from the obstacle grid $A_d$ and the other representing the kernel $K_d$.

If the agent is located near the edge of our known voxel grid we can set the eleemnts of the window would be located in the space beyond the matrix as empty in which case the robot might want to move towards this space or as occupied, which will prevent the robot from moving out of the known grid space.

By employing the Hadamard (element-wise) product (eq.~\ref{eq:hadamard}) between the cutout segment of the obstacle grid \(A_d\) and the corresponding 3D convolutional kernel \(K_d\) for direction \(d\), we derive the resultant matrix \(C_d\), which can be expressed as:

\begin{equation}
	C_d = A_d \odot K_d
	\label{eq:hadamard}
\end{equation}

Subsequently, the avoidance velocity in the Cartesian coordinate system is obtained (eq.~\ref{eq:summation}) by summing all the values in the resulting matrix $C_d$, which can be represented as:

\begin{equation}
	\dot{x}_{poi} = \sum_{i}\sum_{j}\sum_{k} c_{dijk} 
	\label{eq:summation}
\end{equation}

@@

The fundamental concept of our directional kernels lies in computing the repulsive field individually for each direction within the Cartesian coordinate system. \remove{Our filters structure was inspired by the Sobel  operator, a 2D convolutional filter frequently utilized in computer vision for calculating image gradients at specific points.}

Our kernels are designed as three-dimensional matrixes with a primary kernel axis aligned along a specific Cartesian direction, corresponding to the calculated repulsive velocity. The two secondary kernel axes are orthogonal to this primary axis. The distribution of values along the primary axis is inversely symmetric, exhibiting positive values on one side and negative values on the other, with the zero valued cell in the center of the kernel, where jump between max positive and max negative magnitude is. The function of the increase in magnitude along the primary axis of the kernel defines the shape of the repulsive velocity field, determining how the repulsive velocity changes as the agent approaches an obstacle.

The length of the primary axis is critical, as it dictates the detection range for obstacles. Longer kernels can detect obstacles further away
from the robot, essentially extending the ’safety zone’ around the robot. If the primary axis is too long, it can lead to extra calculations and may cause the robot to unnecessarily avoid obstacles that aren't in its immediate path, making its movement and path planning less efficient. A kernel with a primary axis that is too short might restrict the robot's ability to maneuver, detecting obstacles potentially too late, compromising the robots capacity to avoid obstacles effectively (eq. \ref{eq: detect range}). Moreover, it is essential for the magnitudes at the kernel's periphery to be minimal, promoting a smooth increase in repulsive velocity when approaching the obstacle rather than a sudden spike.

\begin{equation}
	\label{eq: detect range}
	num_{primary} = \frac{2 \times \mathrm{range}}{\Delta \mathrm{grid}}\\
\end{equation}

%\remove{In the orthogonal axes, the magnitude of values diminishes with increasing distance from the core of the kernel, nearing zero at the edges. This gradation is vital for a smooth velocity transition when obstacles emerge in the kernels range. The peak magnitude is located at the kernel's core, along the primary axis. The polarity of the values is determined by their position relative to the primary axis. }

The length of the orthogonal axes influences the peripheral detection range for obstacles. Excessively wide kernels may generate repulsive velocities for objects that are not in the path of the robot, whereas too narrow kernels might only detect obstacles aligned directly with the Cartesian direction in the point of interest. When selecting the width and height of the kernel, we must consider the density of the neighboring points of interest on the robot, ensuring that the \replace{collective fields}{combination of kernels} adequately cover the entire manipulator's surrounding area.

Če imamo opravka s točkastim agentom, kot je recimo štirikopet - dron, potem je dobra oblika posameznih matrik enaka dimenzija dolžine, širine in višine. Tako dobimo enakomerno pokritost krogle prostora v okolici agenta in preprečimo mrtve kote, ki se sicer lahko pojavijo, ko se ovira nahaja v bližini agenta, vendar med posameznimi jedri. \add{ADD:primer,slika}

\comment{}{PLOT: kernel 2D images}
\comment{}{EQUATION: kernel generation / values equations}

\add{choice of velocity profile (gaussian, linear ... maybe could move this part to implementation / experiments)}

@@

It is essential that the velocity contributions affecting the robot change smoothly. However, since our obstacle grid is discretely defined, achieving perfect continuity can be challenging. Increasing the resolution of the obstacle field can theoretically bring us closer to continuous behavior, but in practice, we are constrained by finite resolution. To ensure that the velocity remains continuous when transitioning from one cell of the obstacle grid to another at a point of interest (POI), we employ trilinear interpolation. \remove{This technique allows for a smooth and continuous linear approximation of velocities in all three Cartesian directions (x, y, and z) as the POI moves between cells.}

We start by scalling the coordinates of POI into the grid koordinate system, by multiplying it by grid resolution (eq.~\ref{eq:scale grid}).

\begin{equation}
	\label{eq:scale grid}
	\vec{P} = \vec{p}_{\mathrm{POI}} \times\Delta \mathrm{grid}
\end{equation}

We get the indexes of the surrounding cells by first scaling the POI position by grid resolution and than rounding the position to the nearest lower and upper integer positions (eq.~\ref{eq:floor and ceil}).

%\begin{equation}
% X = \left[ \lfloor p_{x} \rfloor, \lceil p_{x} \rceil \right] 
%\end{equation}
%
%\begin{equation}
% Y = \left[ \lfloor p_{y} \rfloor, \lceil p_{y} \rceil \right] 
%\end{equation}
%
% \begin{equation}
% Z = \left[ \lfloor p_{z} \rfloor, \lceil p_{z} \rceil \right] 
%\end{equation}

\begin{equation}
	\label{eq:floor and ceil}
	\vec{P} =
	\begin{bmatrix}
		X \\
		Y \\
		Z \\
	\end{bmatrix}
	=
	\begin{bmatrix}
		\; \lfloor \vec{p}_{\mathrm{POI}}(1) \rfloor \; \lceil \vec{p}_{\mathrm{POI}}(1) \rceil \;  \\
		\; \lfloor \vec{p}_{\mathrm{POI}}(2) \rfloor \; \lceil \vec{p}_{\mathrm{POI}}(2) \rceil \; \\
		\; \lfloor \vec{p}_{\mathrm{POI}}(3) \rfloor \; \lceil \vec{p}_{\mathrm{POI}}(3) \rceil \; 
	\end{bmatrix}
\end{equation}

Once we got the indexes of the eight surrounding cells of our POI, we use our kernel matrix multiplication method, to calculate the 3x1 repulsive velocity vectors for all the cells (eq.~\ref{eq: calc rep vel}).

\begin{equation}
	\label{eq: calc rep vel}
	\vec{V}rep_{xyz,ijk} = \mathrm{calc\_rep\_vel}(X[i], Y[j], Z[k]) \quad \forall i, j, k \in \{1, 2\}
\end{equation}

Trilinear interpolation method works on a 3-dimensional regular grid. Before we can start with the interpolation we need to calculate the distance between POI and smaller coordinates of the cells where we calculated the repulsive velocities (eq.~\ref{eq: deltas interp}). \replace{Since the repulsive values we calculate for the cells are alligned with the centers of the cells, we need to move before the interpolation the positions of known grid points by half of the cell width.}{The calculated repulsive velocity values are located at the centers of the cells. Therefore, before interpolation, we shift the values of the cells coordinates by half the resolution of the obstacle grid for each direction. }

%\begin{equation}
%	\Delta X = \frac{\left( P_x - \left( X(1) + \frac{1}{2} \Delta \mathrm{grid} \right)  \right)}{\left( X(2) - X(1) \right)}
%\end{equation}
%
%\begin{equation}
%	\Delta Y = \frac{\left( P_y - \left( Y(1) + \frac{1}{2} \Delta \mathrm{grid} \right)  \right)}{\left( Y(2) - Y(1) \right)}
%\end{equation}
%
%\begin{equation}
%	\Delta Z = \frac{\left( P_z - \left( Z(1) + \frac{1}{2} \Delta \mathrm{grid} \right)  \right)}{\left( Z(2) - Z(1) \right)}
%\end{equation}

\begin{equation}
	\label{eq: deltas interp}
	\Delta \vec{P} =
	\begin{bmatrix}
		\Delta x \\
		\Delta y \\
		\Delta z		
	\end{bmatrix}
	=
	\begin{bmatrix}
		\frac{\left( P_x - \left( X(1) + \frac{1}{2} \Delta \mathrm{grid} \right)  \right)}{\left( X(2) - X(1) \right)} \\
		\frac{\left( P_y - \left( Y(1) + \frac{1}{2} \Delta \mathrm{grid} \right)  \right)}{\left( Y(2) - Y(1) \right)} \\
		\frac{\left( P_z - \left( Z(1) + \frac{1}{2} \Delta \mathrm{grid} \right)  \right)}{\left( Z(2) - Z(1) \right)} \\
	\end{bmatrix}
\end{equation}

The result of the interpolation is independent of the order of the operations. We first interpolate along the x-axis, followed by along the y-axis and finally along z-axis.

%\begin{equation}
%	\vec{V}_{rep,00} = \vec{V}_{rep,000}(1 - x_d) + \vec{V}_{rep,100}x_d
%\end{equation}
%\begin{equation}
%	\vec{V}_{rep,01} = \vec{V}_{rep,001}(1 - x_d) + \vec{V}_{rep,101}x_d
%\end{equation}
%\begin{equation}
%	\vec{V}_{rep,10} = \vec{V}_{rep,010}(1 - x_d) + \vec{V}_{rep,110}x_d
%\end{equation}
%\begin{equation}
%	\vec{V}_{rep,11} = \vec{V}_{rep,011}(1 - x_d) + \vec{V}_{rep,111}x_d
%\end{equation}

\begin{equation}
	\label{eq: interp x}
	\vec{V}rep_{xyz,jk} = \vec{V}rep_{xyz,0jk}(1 - \Delta x) + \vec{V}rep_{xyz,1jk} \, \Delta x \quad \forall \, j, k \in \{1, 2\}
\end{equation}

%\begin{equation}
%	\vec{V}rep_{xyz,0} = \vec{V}rep_{xyz,00}(1 - \Delta y) + \vec{V}rep_{xyz,10} \, \Delta y
%\end{equation}
%
%\begin{equation}
%	\vec{V}rep_{xyz,1} = \vec{V}rep_{xyz,01}(1 - \Delta y) + \vec{V}rep_{xyz,11} \, \Delta y
%\end{equation}

\begin{equation}
	\label{eq: interp y}
	\vec{V}rep_{xyz,k} = \vec{V}rep_{xyz,0k}(1 - \Delta y) + \vec{V}rep_{xyz,1k} \, \Delta y \quad \forall \, k \in \{1, 2\}
\end{equation}

\begin{equation}
	\label{eq: interp z}
	\vec{V}rep_{xyz} = \vec{V}rep_{xyz,0}(1 - \Delta z) + \vec{V}rep_{xyz,1} \, \Delta z 
\end{equation}

The final result is a repulsive velocity vector that transitions smoothly between the discrete values calculated at distinct points in the obstacle grid.
